Can We Crunch the Social Graph on the Cheap?

### 2011-01-26 doc/hotpar2011/softxmt-hotpar2011.tex
resolved conflicts

\section{Implementation}
\label{sec:approach}

The goal of our implementation is to use multithreading to tolerate
the latency of random accesses to memory in a global address space
spread across multiple nodes. This goal has three key challenges:
\begin{itemize}
\item {\bf Lightweight multithreading}

@@

  In a system with high degrees of concurrency, lightweight
  synchronization primitives are important.
\end{itemize}
We discuss our approach to solving each of these challenges in
turn. For each one, we discuss both our current single-node
implementation and ideas for a multi-node implementation.

@@

Our approach to supporting many lightweight threads uses a cooperative,
user-level multitasking system built using coroutines.

There are three main concerns in designing the coroutine
mechanism. First, the working set of each coroutine must be small, so
many coroutines can be active without trashing the cache. Second,
context switches must be fast, so that overhead does not swamp actual
work. Third, when a coroutine yields, we must choose a suitable
coroutine to run next.

We minimize context size by treating context switches as function
calls, as in~\cite{charm}. This allows us to save and restore the
minimum number of registers allowed by the ABI, and depend on the
compiler to save and restore any other registers in use. We minimize
context switch time both by keeping contexts small and by doing all
switching and scheduling in user space, as in \todo{cite other green
  threads packages, like qthreads}.

There are a number of possibilites for coroutine scheduling.
Ideally, a thread would be rescheduled when its request had
completed, but this may complicate the scheduling decision and slow
down context switches. A simple round-robin scheduler may be sufficient.

\subsubsection{Single-node implementation}
We implemented coroutines as described, with a round-robin
scheduler. If a prefetched data value is no longer in the cache when
the coroutine is rescheduled and the value is used, the coroutine pays
the full cost of the fetch.

\subsubsection{Multi-node ideas}
We will explore how the memory manager and coroutine scheduler may
interact, so that threads waiting on long-latency memory operations or
synchronization operations will be scheduled soon after their
operations have completed.

>>>>

\section{Implementation}
\label{sec:approach}
Our goal is to use multithreading to tolerate
the latency of random accesses to memory in a global address space
spread across multiple nodes.  Our implementation must provide three features:
\begin{itemize}
\item {\bf Lightweight multithreading}

@@

  We must support global synchronization between threads that allows unrelated threads to proceed unblocked.
\end{itemize}
We discuss our approach to solving each of these challenges in
turn. For each one, we discuss both our current single-node
implementation and ideas for a multi-node implementation.

@@

Our approach to supporting many lightweight threads uses a cooperative,
user-level multitasking system built using coroutines. Much work has been done on similar user-level systems \cite{ult,capriccio}, but we have stringent requirements on size and performance, since we have large numbers of contexts rapidly issuing prefetches and passing control.  Hence coroutines must be small, so
many coroutines can be active without trashing the cache, and
context switches must be fast (a small fraction of memory latency) so that we can overlap memory requests and achieve concurrency.

We minimize context size by treating context switches as function
calls, as in~\cite{charm}. This allows us to save and restore the
minimum number of registers allowed by the ABI, and depend on the
compiler to save and restore any other registers in use. We minimize
context switch time by doing all switching and scheduling in user space (kernel traps are at least an order of magnitude slower.)

\subsubsection{Single-node implementation}
We implemented coroutines as described, with a round-robin
scheduler. Coroutines do local work, issue a remote prefetch, and switch to the next in turn.  When the coroutine is reactivated, it assumes the requested data is available and issues a read.  This is a disadvantage of the simple implementation; if the data is not yet received and in cache, the core (and all its coroutines) can stall.

\subsubsection{Multi-node ideas}
We will explore how the memory manager and coroutine scheduler may
interact, so that threads waiting on long-latency memory operations or
synchronization operations will be scheduled soon after their
operations have completed (and not before).
